Symmetric Peer-to-Peer Applications

### 2022-08-22 tpd-22.tex
...

@@

The middleware calls \code{cb\_ini} twice: at the beginning and at the end of
the loop.
It should be used to initialize (and finalize) the network topology, as well
as static globals that live outside the simulation memory, such as the window
and image textures.

{\footnotesize
\begin{verbatim}
void cb_ini (int ini) {
    if (ini) {
        // create the window
        // open arrow images
        // create peer links (peer-id, ip, port)
        p2p_link(2, "192.168.1.2", 5000);
    } else {
        // create the window
        // open arrow images
        // destroy a peer links
        p2p_unlink(2);
    }
}
\end{verbatim}
}

The middleware calls \code{cb\_sim} on every frame to affect the simulation
state \code{G} deterministically.
The callback must never perform any side effects, such as rendering the frame.
The calls normally happen during real-time simulation, i.e., while the user is
interacting with the application.
However, if the peer is out of sync, the middleware may ``time travel'' and
call the simulation many times to go back and forth and catch up with real
time.
This is the reason why \code{cb\_sim} should

This is why it is important that the simulation is detached from the view.

@@

at the same time as the user
is interacting with the application

twice: at the beginning and at the end of
the loop.
It should be used to initialize (and finalize) the network topology, as well
as static globals that live outside the simulation memory, such as the window
and image textures.

The core of the API are the 4 callbacks

The network topology is dynamic

@@

The middleware relies on the library \emph{SDL} to deal with the network and
input events, such as timers, mouse clicks and key presses.
The code that follows is a typical:

{\footnotesize
\begin{verbatim}
// preallocated full application state
struct {
    ...
} G;

// executes a simulation step (handles ticks or events)
void cb_sim (p2p_evt evt) {
    switch (evt.id) {
        ...
    }
}

// executes the simulation effects (e.g., redraw the screen)
void cb_eff (int trv) {
    ...
}

// generates events
int cb_rec (SDL_Event* sdl, p2p_evt* evt) {
    switch (sdl->type) {
    	...
    }
}

\end{verbatim}
}

for computer graphics, and
\emph{Arduino} for embedded systems.%
\footnote{SDL: \url{www.libsdl.org}. Arduino: \url{www.arduino.cc}.}
This allows for an easier integration of our middleware with practical systems,
and also reinforces how, in our proposal, programming distributed versions
becomes similar to their local counterparts.

{\footnotesize
\begin{verbatim}
void p2p_init (
    uint8_t me, int port, int fps,
    int mem_n, void* mam_app,
    void (*cb_sim) (p2p_evt),           // simulation callback
    void (*cb_eff) (int trv),           // effects callback
    int (*cb_rec) (SDL_Event*,p2p_evt*) // event recording callback
);
void p2p_link (char* host, int port, uint8_t me);
void p2p_quit (void);
void p2p_loop (void);

void p2p_bcast (uint32_t tick, p2p_evt* evt);
void p2p_dump  (void);

 > freechains '#p2p.md' dislike 3_AE3A1B.. --sign=USR1
 > freechains '#p2p.md' dislike 3_AE3A1B.. --sign=USR2
 > freechains '#p2p.md' dislike 3_AE3A1B.. --sign=USR3
 > freechains-vcs '#p2p.md' checkout p2p.md
 > cat p2p.md
 P2p networking is...
 The [USENET](#usenet.md), ...  <-- no 3rd line
\end{verbatim}
}

>>>>

Under the hood, the middleware relies on the library \emph{SDL}%
\footnote{SDL: \url{www.libsdl.org}}
to deal with the network and input events, such as timers, mouse clicks and
key presses.

@@

The middleware calls \code{cb\_ini} twice: at the beginning and at the end of
the loop.
It should be used to initialize (and finalize) the network topology, as well
as static globals that can live outside the simulation memory, such as the
window and image textures.

{\footnotesize
\begin{verbatim}
void cb_ini (int ini) {
    if (ini) {
        // create the SDL window
        // open the arrow PNG images
        // create peer links (peer-id, ip, port)
        p2p_link(2, "192.168.1.2", 5000);
    } else {
        // destroy the SDL window
        // destroy the arrow PNG images
        // destroy peer links
        p2p_unlink(2);
    }
}
\end{verbatim}
}

The middleware calls \code{cb\_sim} on every frame to affect the simulation
state of \code{G} deterministically.
The callback must never perform any side effects, such as stateful calls or
rendering the frame.
The calls normally happen during real-time simulation, i.e., while the user is
interacting with the application.
However, if the peer is out of sync, the middleware may ``time travel'' and
call the simulation many times to go back and forth and catch up with real
time.
This is the reason why \code{cb\_sim} should be split from the rendering
callback \code{cb\_ren}~\cite{tml.js}.
%, which the middleware calls only on real-time frames.

cbsim

In our example, this callback just needs to redraw the rectangle at the
current position:

{\footnotesize
\begin{verbatim}
void cb_eff (void) {
    // clears the SDL window
    // redraws the rectangle at the current position
    SDL_Rect r = { G.x, G.y, 20, 20 };
    SDL_DrawRect(&r);
}
\end{verbatim}
}

Finally, the \code{cb\_evt} callback is called by the middleware in real time,
whenever a local SDL event occurs.
This callback has two goals:
    (i)  broadcast application events, and
    (ii) provide instant feedback to the user.
Not all local low-level events may generate application events to be broadcast
in the network.
The callback is allowed to perform side effects, such as modifying the network
topology, or exhibiting some feedback on the screen.
In our example, we only generate application events for the 5 key events, and
we also exhibit the pressed key in real time:

{\footnotesize
\begin{verbatim}
int cb_evt (SDL_Event* sdl, p2p_evt* evt) {
    if (sdl->type == SDL_KEYDOWN) {
        int key = sdl->key.keysym.sym;
	if (key == <any-of-the-keys>) {
            SDL_DrawImage(<img-of-the-key>);
            *evt = (p2p_evt) { P2P_EVT_KEY, key };
            return 1; // generate application event
	}
    }
    return 0; // otherwise, does not generate any event
}
\end{verbatim}
}

@@

- p2p links por fora, como um servico TCP local
- The network topology is dynamic

@@
